Prediction of Protein-Protein Interactions on the Human and Rice Interactomes

### 2021-01-22 nicolas/PPIPL3/bmc_template/framework.tex
updated document structure and script

For each metric, the initial prediction in step A is used to train the supervised 
ML model

@@

However, the interest of this study is to assess if Machine Learning can boost
the overall performance on the prediction, so the proposed procedure
is adapted to the human interactome data: Instead of having to remove
a fraction of edges from the network in order to predict and validate,
the prediction network and the validation networks are given beforehand.
The rest of the procedure is carried out as mentioned above.
Table \ref{T1} presents the results for the combinations of Node2Vec
with each feature.

>>>>

First, for each metric, the initial prediction in stage A is used to train the 
supervised ML model. It is important to keep in mind that the training dataset 
and the validation dataset should be sampled randomly to ensure an even
distribution of the labeling. However, the intrinsic imbalance in the
datasets has been addressed with under-sampling, so that the training 
dataset (80\% of data) has the same amount of interactions labeled true and false. 
Second, the validation dataset (20\% of data) is introduced in the model and it returns 
a score in the range $(0,1)$ assessing the probability that an edge is feasible. 
The considered metric to evaluate the ML model performance is the area under the 
the curve (AUC) of the Receiver Operating Characteristic (ROC). This curve
corresponds to plot the sensitivity (probability of predicting a real
positive as positive) against 1-specificity (probability of predicting
a real negative as positive). AUC values of 1 represent a perfect
prediction and 0.5 corresponds to a random guess. Normally, values
over 0.8 of AUC are considered good.

@@

However, the interest of this study is to assess if Machine Learning can boost
the overall performance on the prediction, so the proposed procedure 
is adapted to the human interactome data: the prediction network and 
the validation networks are given beforehand, so that training is done using 
the labels from one side and the validation is done with the other side. 
The rest of the procedure is carried out as mentioned above.
Table \ref{T1} presents the results for the combinations of Node2Vec
with each feature.

### 2021-01-22 nicolas/PPIPL3/bmc_template/preliminaries.tex
updated document structure and script

% "Preliminaries" refers to fundaments for the new reader on the main topics of
% the paper

Let $G=(V,E), n=|V|, m=|E|$ be an undirected graph describing $n$
proteins and $m$ pairwise interactions between them. $G$ can be represented as a Boolean 
adjacency matrix $A$ of size $n\cdot n$, where the value for a position
$(i,j), 0\leq i,j < n$ is set to $1$ if there is a direct interaction
between proteins $i$ and $j$, or set to $0$ otherwise. As it is an undirected network, $A$ is symmetric.

@@

>>>>

% "Preliminaries" refers to fundaments for the new reader on the main topics of
% the paper
\subsection*{Protein Interactions Networks}
Let $G=(V,E)$ be an undirected graph describing $n=|V|$ proteins as 
nodes and $m=|E|$ pairwise interactions between the graph as edges. 
$G$ can be represented as a Boolean adjacency matrix $A_G$ of size 
$n^2$, where the value of the entry $(i,j), 0\leq i,j < n$ is set 
to $1$ if there is a direct interaction between proteins $i$ and $j$ 
and set to $0$ otherwise. As it is an undirected network, $A_G$ is symmetric.

@@

\subsection*{Feature Learning of a Network and node2vec}
The problem of feature learning in a network can be addressed as a maximum 
likelihood optimization problem, in which the function (to be learned) 
$f : V \rightarrow \mathbb{R}^d$ maps from nodes to d-dimensional feature 
representations. $f$ can also be seen as a matrix of size $|V| \times d$. 
The optimization problem is complex but tractable when assuming conditional 
independence and spatial symmetry in feature space. Namely, conditional 
independence means that the likelihood of observing a neighbor node does 
not depend on any other neighbor node.

For every node $u \in V$, there exists a network neighborhood, denoted as 
$N_S (u) \subset V$, generated by a sampling strategy limited to k to be 
able to compare different strategies fairly. When sampling, one possible strategy is to retrieve nodes as close as possible from the source. This strategy is known as Breadth-first Sampling (BFS). On the other hand, one can move as far as possible from the source, also known as Depth-first Sampling (DFS). In this regard, node2vec generates random walks from each node with user-defined weighting for back-edges (return parameter p) and for edges that go away from source (in-out parameter q). Based on the information (paths) gathered from these walks, node2vec uses a semi-supervised method to learn rich feature representations for nodes in a network.

Since the purpose of this study is link prediction, another mapping function is defined for the edges, so that its feature representation is a bootstrap over the feature representation of nodes. Let $g : V \times V \rightarrow \mathbb{R}^d$, then the function $g$ is defined for any pair of nodes $(u,v)$, even if this edge does not exist in the network. For this work, the Hadamard operator is used, which consists of an element-wise product of the node features.

\subsection*{Gradient Boosted Trees and XGBoost}
(THIS IS PENDING)
